Towards Bridging the Gap... [TBD]

### 2014-03-14 paper/draft.tex
Merge
Added Related Works paragraph

%%% PAGE DIMENSIONS
%\usepackage{geometry} % to change the page dimensions
%\geometry{a4paper} % or letterpaper (US) or a5paper or....
% \geometry{margin=2in} % for example, change the margins to 2 inches all round
% \geometry{landscape} % set up the page for landscape
%   read geometry.pdf for detailed page layout information
%\usepackage[hidelinks]{hyperref}
%\usepackage[autostyle,french=guillemets,babel]{csquotes}
\usepackage{graphicx} % support the \includegraphics command and options
%\usepackage[backend=biber]{biblatex}
%\addbibresource{biblio.bib}
% \usepackage[parfill]{parskip} % Activate to begin paragraphs with an empty line rather than an indent
%%% PACKAGES
\usepackage{booktabs} % for much better looking tables
\usepackage{array} %
% for better arrays (eg matrices) in maths
\usepackage{paralist} % very flexible & customisable lists (eg. enumerate/itemize, etc.)
\usepackage{verbatim} % adds environment for commenting out blocks of text & for better verbatim
%\usepackage{subfig} % make it possible to include more than one captioned figure/table in a single float
% These packages are all incorporated in the memoir class to one degree or another...
\usepackage{amsfonts}

@@

\title{Towards Bridging the Gap... [TBD]}

\author{
% You can go ahead and credit any number of authors here,
% e.g. one 'row of three' or two rows (consisting of one row of three
% and a second row of one, two or three).
%
% The command \alignauthor (no curly braces needed) should
% precede each author name, affiliation/snail-mail address and
% e-mail address. Additionally, tag each line of
% affiliation/address with \affaddr, and tag the
% e-mail address with \email.
%
% 1st. author
\numberofauthors{3}
\alignauthor
Alex Palesandro\\
\affaddr{Orange Labs, France}\\
\texttt{palexster@gmail.com}
\alignauthor
Aurélien Wailly\\
\affaddr{Orange Labs, France}\\
\texttt{aurelien.wailly@orange.com}
\alignauthor
Marc Lacoste\\
\affaddr{Orange Labs, France}
\texttt{marc.lacoste@orange.com}
}
%\setlist[itemize]{leftmargin=*}
\begin{document}
\maketitle

@@

To resume, in order to present a complete view of the architecture, we identified five key design principles, that a cloud computing platform has to implement in order to solve the cited problems. In the following paragraphs, those principles will represent a grid of analysis of different architectures.
\begin{itemize}[]
\item \textbf{DP1 User-Data privacy} A curious or malicious administrator ( or another platform user) would not have the right to inspect and analyze user data and instances, without the explicit consensus with the user above mutually trusted services. User-data privacy is a major challenge of modern cloud platforms. It is important to identify a trade-off that allows providers to identify and intervene on possible menaces, but also guarantee the impossibility for a curious or malicious operator to access to them without restrictions. In order to respect this principle when a platform administrator manages the platform it has not the highest level of privilege and each action on users object that he wants to deliver has to be validated by a mutually-trusted core, executed at the highest level of privilege. Moreover, this mutually-trust core has also to apply memory page cipher/uncipher in order to prevent memory page snooping when a page is swapped out.
\item \textbf{DP2 Fail-Safe modular architecture} In order to recover soft failures, the archtecure is supposed to be fail-safe, being able of automatically migrate instances or user-data in presence of a menace, but also reactivating the correct functioning in presence of a software failure. In order to satisfy this property, the amount of safety-critical components has to be as lowest as possible and the overall design has to be kept as modular as possible
\item \textbf{DP3 Small TCB} The amount of lines of code executed at the highest level of privilege is normally proportional to the number of potential failures and bugs in the platform. According to this, an architecture presenting a tiny TCB will improve its safety and integrity properties by design.
\item \textbf{DP4 Inter-platform support} The cloud architecture has to be designed in order to support directly different platforms and service providers. The same user virtual instances have to be executed ( and potentially migrated) unmodified over different platforms.
\item \textbf{DP5 3rd part compatibility} Avoid disruption of existing 3rd part applications compatibility could represent a major principle to consider. Platform interfaces has to be kept as close as possible to the existing one, preventing application rewrite.
\end{itemize}
Moreover, beside this functional design principles specifying which features have to be achieved by the platform, another set of non-design principles could be identified. Those principles does not concern directly features or aspects of the architecture of the platform, but the effective feasibility of the architecture. 
\begin{itemize}[]
\item \textbf{NDP1 Minimize code addition} a major issue could be represented by the amount of code lines that have to be added and maintained to an existing architecture to meet the functional requirement. Even it does not represent strictly a technical limitation to a certain proposal, it could become a ``show-stopper'' factor for its utilization.
\end{itemize}
In our analysis, it will be kept in consideration not only the overall suitability of the architecture, but also a quantification of the effort needed to develop and deploy it. In table \ref{int:des} it is possible to observe a resume of previous cited principles.

@@

% Perfomance issuess adressed
However, several works demonstrate that it is possible to put in place low overhead nested virtualization using special approaches \cite{turtle:ibm} . Moreover, other works try to recover performance loss exploiting multi-level consolidation, introduced in nested virtualization \cite{art:blan}. In conclusion, nested virtualization slowdown could be considered at now an acceptable price to achieve advanced features and, in addition, hardware manifacturers are developing and proposing new hardware facilities to reduce this overhead \cite{vmcs:nakajima}.

@@

@@

Finally, the above architecture does not improve the overall stability of the infrastructure. The underlying core, L0, has the same problem of single-layer GPH architecture: the large TCB implies also that a failure experienced in one of the critical components will simply produce a generalized failure of the whole platform, not satysfing DP2. Moreover, in GPH standard architecture, the administrator has the entire control of the platform and could potentially snoop and inspect user VM status and data.

@@

In addition, MH inherits the concept of micro-kernel to implement different ``personalities''. Those ``personalities'' are concretely a set of ``processes'' handling specific functions running in user-space. As those functions are implemented outside the kernel, they are not unique but they are changeable and could coexist at the same time. This feature was seen with interest because of the possibility to move some virtualization-related from the kernel space to user-space and to strictly separate piece of software executed for different users. This extraction process drives to a more solid platform, considering that user-space crashes are generally not fatal for the entire system integrity.  A possible example, that will be analyzed in next section, is the user-space VMM component of Nova, Vancouver, that resides in user-space \cite{nova}.

@@

@@

\subsection{GPH over MH}
\label{par:gom}

@@

\begin{figure}
\begin{center}
\includegraphics[width=0.6\columnwidth]{article.png}
\end{center}
\caption{Xen over NOVA architecture}
\label{fig:io}
\end{figure}

>>>>

%%% PAGE DIMENSIONS
%\usepackage{geometry} % to change the page dimensions
%\geometry{a4paper} % or letterpaper (US) or a5paper or....
% \geometry{margin=2in} % for example, change the margins to 2 inches all round
% \geometry{landscape} % set up the page for landscape
%   read geometry.pdf for detailed page layout information
%\usepackage[hidelinks]{hyperref}
%\usepackage[autostyle,french=guillemets,babel]{csquotes}
\usepackage{graphicx} % support the \includegraphics command and options
\usepackage[backend=biber]{biblatex}
\addbibresource{biblio.bib}
% \usepackage[parfill]{parskip} % Activate to begin paragraphs with an empty line rather than an indent
%%% PACKAGES
\usepackage{booktabs} % for much better looking tables
\usepackage{array} %
% for better arrays (eg matrices) in maths
\usepackage{paralist} % very flexible & customisable lists (eg. enumerate/itemize, etc.)
\usepackage{verbatim} % adds environment for commenting out blocks of text & for better verbatim
%\usepackage{subfig} % make it possible to include more than one captioned figure/table in a single float
% These packages are all incorporated in the memoir class to one degree or another...
\usepackage{amsfonts}

@@

\title{}

\author{
% You can go ahead and credit any number of authors here,
% e.g. one 'row of three' or two rows (consisting of one row of three
% and a second row of one, two or three).
%
% The command \alignauthor (no curly braces needed) should
% precede each author name, affiliation/snail-mail address and
% e-mail address. Additionally, tag each line of
% affiliation/address with \affaddr, and tag the
% e-mail address with \email.
%
% 1st. author
\numberofauthors{3}
\alignauthor
Alex Palesandro\\
\affaddr{Orange Labs, France}\\
\texttt{alex.palesandro@gmail.com}
\alignauthor
Aurélien Wailly\\
\affaddr{Orange Labs, France}\\
\texttt{aurelien.wailly@orange.com}
\alignauthor
Marc Lacoste\\
\affaddr{Orange Labs, France}
\texttt{marc.lacoste@orange.com}
}
%\setlist[itemize]{leftmargin=*}
\begin{document}
\maketitle

@@

To resume, in order to present a complete view of the architecture, we identified five key design principles, that a cloud computing platform has to implement in order to solve the cited problems. In the following paragraphs, those principles will represent a grid of analysis of different architectures.
\begin{itemize}[]
\item \textbf{DP1 User-Data privacy} A curious or malicious administrator ( or another platform user) would not have the right to inspect and analyze user data and instances, without the explicit consensus with the user above mutually trusted services. User-data privacy is a major challenge of modern cloud platforms. It is important to identify a trade-off that allows providers to identify and intervene on possible menaces, but also guarantee the impossibility for a curious or malicious operator to access to them without restrictions. In order to respect this principle when a platform administrator manages the platform it has not the highest level of privilege and each action on users object that he wants to deliver has to be validated by a mutually-trusted core, executed at the highest level of privilege. Moreover, this mutually-trust core has also to apply memory page cipher/uncipher in order to prevent memory page snooping when a page is swapped out.
\item \textbf{DP2 Fail-Safe modular architecture} In order to recover soft failures, the archtecure is supposed to be fail-safe, being able of automatically migrate instances or user-data in presence of a menace, but also reactivating the correct functioning in presence of a software failure. In order to satisfy this property, the amount of safety-critical components has to be as lowest as possible and the overall design has to be kept as modular as possible
\item \textbf{DP3 Small TCB} The amount of lines of code executed at the highest level of privilege is normally proportional to the number of potential failures and bugs in the platform. According to this, an architecture presenting a tiny TCB will improve its safety and integrity properties by design.
\item \textbf{DP4 Inter-platform support} The cloud architecture has to be designed in order to support directly different platforms and service providers. The same user virtual instances have to be executed (and potentially migrated) unmodified over different platforms.
\item \textbf{DP5 3rd part compatibility} Avoid disruption of existing 3rd part applications compatibility could represent a major principle to consider. Platform interfaces has to be kept as close as possible to the existing one, preventing application rewrite.
\end{itemize}
Moreover, beside this functional design principles specifying which features have to be achieved by the platform, another set of non-design principles could be identified. Those principles does not concern directly features or aspects of the architecture of the platform, but the effective feasibility of the architecture. 
\begin{itemize}[]
\item \textbf{NDP1 Minimize code addition} a major issue could be represented by the amount of code lines that have to be added and maintained to an existing architecture to meet the functional requirement. Even it does not represent strictly a technical limitation to a certain proposal, it could become a ``show-stopper'' factor for its utilization.
\end{itemize}
In our analysis, it will be kept in consideration not only the overall suitability of the architecture, but also a quantification of the effort needed to develop and deploy it. In table \ref{int:des} it is possible to observe a resume of previous cited principles.

@@

% Perfomance issuess adressed
However, several works demonstrate that it is possible to put in place low overhead nested virtualization using special approaches \cite{turtle:ibm} . Moreover, other works try to recover performance loss exploiting multi-level consolidation, introduced in nested virtualization \cite{art:blan}. In conclusion, nested virtualization slowdown could be considered nowadays an acceptable price to achieve advanced features and, in addition, hardware manifacturers are developing and proposing new hardware facilities to reduce this overhead \cite{vmcs:nakajima}.

@@

Furthermore, DeHype \cite{wu2013TamHosHypMosDepExe} presented an dissected version of KVM hypervisor, trying to reduce the TCB of this popular hypervisor. The approach consisted in demoting a great part of the hypervisor codebase to a user-level execution. DeHype satisfies all vertical design principles listed before (DS1,DS2,DS3), reinterpreting the microkernel approach of  demoting the greatest part of code to userspace. DeHype increases a lot the robustness of a cloud platform leveraging KVM, limiting for example the possibility of attack in multi-tenancy platform, but it does not address horizontal design principles about a multi-provider service.

@@

Finally, the above architecture does not improve the overall stability of the infrastructure. The underlying core, L0, has the same problem of single-layer GPH architecture: the large TCB implies also that a failure experienced in one of the critical components will simply produce a generalized failure of the whole platform, not satisfying DP2. Moreover, in GPH standard architecture, the administrator has the entire control of the platform and could potentially snoop and inspect user VM status and data.

@@

In addition, MH inherits the concept of micro-kernel to implement different ``personalities''. Those ``personalities'' are concretely a set of ``processes'' handling specific functions running in user-space. As those functions are implemented outside the kernel, each user could rely on a different instances of the hypervisor user-space process, that could coexist at the same time. This process separation between user could enforce a straightforward mechanism to neutralize VM-to-VM attack from a malicious user. Moreover, user-space processes are easier to debug with commodity tools (GDB, valgrind).
Finally, this extraction process drives to a more solid platform, considering that user-space crashes are generally not fatal for the entire system integrity.  A possible example, that will be analyzed in next section, is the user-space VMM component of Nova, Vancouver, that resides in user-space \cite{nova}.
In addition, this approach is not far from the one proposed in DeHype \cite{wu2013TamHosHypMosDepExe}, where a consistent part of the KVM hypervisor has been demoted in user-space.

@@

\begin{figure*}
\begin{center}
\includegraphics[width=2\columnwidth]{globalcut.png}
\end{center}
\caption{}
\label{fig:io}
\end{figure*}

@@

\subsection{GPH over MH}

\label{par:gom}

@@
